### 2020-02-08 presentations/jsps_2020-03-01/index.Rmd
Update slides

# Load targets
loadd(list = c(
  "taxon_id_map",
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "occ_point_data_raw",
  "ja_bioregions_ferns",
  "biodiv_results_ferns",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns"), 
  cache = ja_fern_cache)

@@

```{r alpha-div-ferns-prep}
# Add ecostructure motif scores and canape results to alpha diversity,
# filter out a single location south of 21 lat that had no ferns
alpha_div_ferns <-
  species_motifs_ferns_6L$omega %>%
  as.data.frame %>%
  rownames_to_column("site") %>%
  as_tibble %>%
  separate(site, c("longitude", "latitude"), sep = "_") %>%
  left_join(
    mutate_at(alpha_div_ferns, vars(longitude, latitude), as.character),
    by = c("longitude", "latitude")) %>%
  left_join(
    mutate_at(biodiv_results_ferns, vars(longitude, latitude), as.character),
    by = c("longitude", "latitude")) %>%
  mutate_at(vars(longitude, latitude), as.numeric) %>%
  rename_at(vars(1:6 %>% as.character), ~paste0("motif_", .)) %>%
  filter(latitude > 21)
```

@@

## Biodiversity is in danger from climate change

@@

## We must understand the distribution of biodiversity to save it

- Most analyses of BD only include richness
- But BD is more complicated: evo history also important

@@

## Japan is an ideal site to study fern BD

- Extremely well sampled

@@

## Goals

- Map biodiversity (species richness, PD, endemism)
- Identify biogeographical realms
- Understand drivers of biogeo structure

@@

## Data

- Occurrence data of Japanese fern specimens
- Phylogenetic tree (*rbcL*)
- Morphological traits
- Climate data

@@

## Species richness peaks on Yakushima Isl.

@@

@@

## Phylogenetic diversity is high in N and S

@@

@@

## S islands have high rates of endemism

```{r canape-plot, out.width = "120%", fig.asp = 0.7}
canape_cols <- RColorBrewer::brewer.pal(4, "Dark2") %>%
  set_names(c("mixed", "neo", "palaeo", "super"))

ggplot(biodiv_results_ferns, aes(x = longitude, y = latitude, fill = canape)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_manual(values = canape_cols, na.value = "grey80") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Endemism") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

## Neoendemic areas have more sexual diploids

@@

.center[
```{r canape-sex, fig.asp = 0.7}

percent_sex_dip_ferns_round <-
  percent_sex_dip_ferns %>% 
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3))

canape_sex_mode <-
  biodiv_results_ferns %>%
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3)) %>%
  inner_join(
    percent_sex_dip_ferns_round,
    by = c("latitude", "longitude")
  ) %>%
  assert(is_uniq, secondary_grid_code) %>%
  filter(!is.na(canape))

# Run analysis of variance (AOV) on
# latitudinal breadth by reproductive mode.
canape_by_repro_model = aov(
  percent_sex_dip ~ canape,
  data = canape_sex_mode)

canape_by_repro_model_summary = tidy(canape_by_repro_model)

# AOV of latidudinal breadth by reproductive mode
# showed a significant difference, so
# run Tukey HSD test on results.
canape_by_repro_tukey = TukeyHSD(canape_by_repro_model) %>% tidy

canape_sex_mode %>%
  group_by(canape) %>%
  summarize(
    mean_sex_dip = mean(percent_sex_dip, na.rm = TRUE),
    sd_sex_dip = sd(percent_sex_dip, na.rm = TRUE)
  ) %>%
  mutate(
    upper = mean_sex_dip + sd_sex_dip,
    lower = mean_sex_dip - sd_sex_dip
  ) %>%
  mutate(canape = fct_reorder(canape, mean_sex_dip)) %>%
  ggplot(aes(x = canape, y = mean_sex_dip, fill = canape)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = canape_cols, na.value = "grey80", guide = FALSE) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25) +
  labs(
    x = "Type of endemicity",
    y = "% sexual diploids"
  ) +
  jntools::standard_theme()

@@

@@

## Ecological structure of Japanese ferns, *K* = 6

@@

## Motif 2 tends to cluster, motif 6 tends to overdisperse

>>>>

# Load targets
loadd(list = c(
  "taxon_id_map",
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "occ_point_data_raw",
  "ja_bioregions_ferns",
  "biodiv_results_ferns",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns",
  "repro_data"),
  cache = ja_fern_cache)

@@

```{r alpha-div-ferns-prep}
# Add ecostructure motif scores and canape results to alpha diversity,
# filter out a single location south of 21 lat that had no ferns
alpha_div_ferns <-
  species_motifs_ferns_6L$omega %>%
  as.data.frame %>%
  rownames_to_column("site") %>%
  as_tibble %>%
  separate(site, c("longitude", "latitude"), sep = "_") %>%
  left_join(
    mutate_at(alpha_div_ferns, vars(longitude, latitude), as.character),
    by = c("longitude", "latitude")) %>%
  left_join(
    mutate_at(biodiv_results_ferns, vars(longitude, latitude), as.character),
    by = c("longitude", "latitude")) %>%
  mutate_at(vars(longitude, latitude), as.numeric) %>%
  # Make canape 'NA' an explicit factor level
  mutate(canape = forcats::fct_explicit_na(canape, "NA")) %>%
  rename_at(vars(1:6 %>% as.character), ~paste0("motif_", .)) %>%
  filter(latitude > 21)
```

@@

## 生物多様性は危機にさらされている

- 地球温暖化をはじめ、人間の活動によって生物多様性が失われている

- 保全対策が必要

@@

## 生物多様性の保全には、まずその分布を理解する必要がある

- 今まで生物多様性は主に種数だけで計れてきた

- しかし、生物多様性には様々な側面（次元）がある

 - 系統的多様性

 - 機能的多様性

 - など

@@

## 日本のシダ植物

.pull-left[
- どの国よりも豊富なサンプリング

- 日本シダの会の動力の甲斐

- 生物多様性の多次元解析の最適？
]

.pull-right[
```{r gbif-colls-per-country}
# Read in pteridophyte GBIF collections per country data from
# Iwasaki lab presentation
collections_by_country <- read_csv(here::here("data_raw/gbif_pterido_coll_by_country.csv")) %>%
  mutate(country = factor(country)) %>%
  mutate(country = fct_reorder(country, n))

# Plot showing top 10
ggplot(collections_by_country, aes(x = country, y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "GBIF pteridophyte \ncollections (top 10)",
    y = "No. collections") +
  scale_y_continuous(labels = scales::comma) +
  ggplot2::theme_bw(base_size = 24) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    panel.grid.major.y = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(colour="black"),
      axis.text.y = ggplot2::element_text(colour="black")
  )
```
]

@@

## 日本のシダ植物

.pull-left[
- どの国よりも豊富なサンプリング

- 日本シダの会の動力の甲斐

- 生物多様性の多次元解析の最適？
]

.pull-right[
```{r ja-pterido-coll-curve}
# Format pteridophyte abundances (no hybrids) as numeric vector
pterido_abun <-
occ_point_data_raw %>%
  group_by(species) %>%
  count(sort = TRUE) %>%
  pull(n)

# Run iNEXT
inext_out <- iNEXT::iNEXT(pterido_abun, q=0, datatype="abundance")

# create subset of observed points to add to plot as dots
observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]

# Make iNEXT plot
inext_out$iNextEst %>%
  filter(m < 500000) %>%
  ggplot(aes(x=m, y=qD)) +
  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70", alpha=0.2) +
  geom_line(aes(linetype=method), size=0.8) +
  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
  geom_point(data=observed_data, size=5) +
  scale_y_continuous("No. spp.") +
  scale_x_continuous("No. collections") +
  labs(
    title = "Species collection curve",
    subtitle = "Japanese pteridophytes (no hybrids)"
  ) +
  ggplot2::theme_bw(base_size = 24) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_text(colour="black"),
    axis.text.y = ggplot2::element_text(colour="black"),
    legend.position="bottom", 
    legend.title=element_blank()
    )

@@

## 本研究の目的

**日本産シダ類** において

- 多様性を複数の次元から把握すること

- 生態的構造を検討すること

- 以上のパターンを生み出している要素を探ること

```{r calc-species-num}
# Count number of fern species, no hybrids

# Inner-join fern occurrence data (incl. hybrids)
# with repro data (no hybrids)
n_ferns_no_hybrids <-
occ_data_ferns %>%
  select(taxon_id) %>%
  unique %>%
  inner_join(repro_data) %>%
  pull(taxon_name) %>%
  n_distinct()

## データ

- 日本産シダ類の採取データ

  - `r nrow(occ_data_ferns) %>% scales::comma()`点

  - `r n_ferns_no_hybrids` 種類（雑種を除く）

- 系統樹 (*rbcL*)

- 形態的特徴

- 気候データ

.footnote[Ebihara and Nitta. 2019. *Journal of Plant Research* 132:723–738]

@@

## 屋久島は種数の頂点

@@

## 系統的多様性

- Mean phylogenetic distance (MPD): ある地点に生育する種の**平均的系統距離**

- 生値をランダムな分布と比較して、ランダムより高いか、低いか、解析する

.footnote[Webb, et al. 2002. *Annual Review of Ecology and Systematics* 33:475–505.]

@@

## 系統的多様性は南北に高い

@@

a + b + plot_layout(widths = c(3,1))
```
---
class: big-margin

## Categorical Analyses of Neo- and Paleo-Endemism (CANAPE)

- 分布域が狭く、系統樹の枝が短い：Neoendemic

- 分布域が狭く、系統樹の枝が長い：Paleoendemic

- 両方混じっている地域：Mixed

- 生値をランダムな分布と比較して、ランダムと有意義な差があれば**固有地域**として認める

.footnote[Mishler, et al. 2014. *Nature Communications* 5:4473]

@@

## 南の島々は固有種が多い

```{r canape-plot, out.width = "120%", fig.asp = 0.7}
canape_cols <- RColorBrewer::brewer.pal(4, "Set1") %>%
  c(., "grey80") %>%
  set_names(c("neo", "palaeo", "super", "mixed", "NA"))

biodiv_results_ferns %>%
  mutate(canape = fct_relevel(canape, "palaeo", "mixed", "neo")) %>%
ggplot(aes(x = longitude, y = latitude, fill = canape)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_manual(values = canape_cols, na.value = "grey80") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Endemism") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

## Neoendemicな地域は有性生殖・二倍体種が多い

@@

.center[
```{r canape-sex, fig.asp = 0.7}
# Run analysis of variance (AOV) on
# endemism type by reproductive mode.
canape_by_repro_model <- aov(
  percent_sex_dip ~ canape,
  data = alpha_div_ferns)

# Inspect results:

# - p-value
# tidy(canape_by_repro_model)

# - AOV of endemism type by reproductive mode
# showed a significant difference, so
# run Tukey HSD test on results.
# TukeyHSD(canape_by_repro_model) %>% tidy

# Assign tukey groups for plotting
tukey_groups <-
    multcomp::glht(
    canape_by_repro_model, 
    linfct = multcomp::mcp(canape = "Tukey")) %>%
    multcomp::cld() %>%
    purrr::pluck("mcletters", "Letters") %>%
    tibble(canape = names(.), group = .) %>%
    mutate(
      y_pos = 1.1
    )

alpha_div_ferns %>%
  group_by(canape) %>%
  summarize(
    mean_sex_dip = mean(percent_sex_dip, na.rm = TRUE),
    sd_sex_dip = sd(percent_sex_dip, na.rm = TRUE)
  ) %>%
  mutate(
    upper = mean_sex_dip + sd_sex_dip,
    lower = mean_sex_dip - sd_sex_dip
  ) %>%
  mutate(canape = fct_relevel(canape, "palaeo", "mixed", "neo")) %>%
  ggplot(aes(x = canape, y = mean_sex_dip, fill = canape)) +
  geom_col() +
  geom_text(data = tukey_groups, aes(y = y_pos, label = group), color = "black") +
  scale_y_continuous(labels = scales::percent, breaks = c(0, 0.25, 0.5, 0.75, 1.0)) +
  scale_fill_manual(values = canape_cols, na.value = "grey80", guide = FALSE) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25) +
  labs(
    x = "Type of endemicity",
    y = "% sexual diploids"
  ) +
  jntools::standard_theme()

@@

.footnote['a', 'b', 'c' = Tukey's HSDテストによる有意義な差があったグループ]

## 生態的構造

STRUCTURE (Pritchard 2000)と同じようなアルゴリズムで生態的構造を解析する。

STRUCTUREでは、**遺伝子型**のデータから**個体レベル**の遺伝的構造を推定する

Ecostructure (White 2019)では、**種の分布**のデータから**群集レベル**の生態的構造を推定する

.footnote[Pritchard, et al. 2000. *Genetics* 155: 945-959<br>
White, et al. 2019. *Nature Communications* 10:2646.]

@@

## 日本産シダ類の生態的構造 (*K* = 6)

@@

## 生態的構造のグループによって、系統的多様性が違う
